The Cluster Difference Imaging Photometric Survey (CDIPS). I. Method, Light Curves, and Planet Candidates from Sectors 6--8

### 2019-05-29 paper/ms.tex
did astrometry
paper: update figures and text

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.7\textwidth]{f1_PLACEHOLDER.pdf}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf There are many clusters; TESS looks at them.} Placeholder skymap.
		\label{fig:clustermap}
	}
\end{figure*}

@@

\begin{figure*}[!ht]
  \gridline{\fig{f5a.png}{0.95\textwidth}{}}
  \vspace{-0.8cm}
  \gridline{\fig{f5b.png}{0.95\textwidth}{}}
  \caption{
      {\it Top.} Cross-match statistics from \cite{Kharchenko_et_al_2013} {\it 
      vs.} Gaia-DR2.
      {\it Bottom.} Ditto, for \cite{dias_proper_2014} {\it vs.} Gaia-DR2.
  }
  \label{fig:xmatch_info}
\end{figure*}

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.3\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.2cm}
	\caption{
    {\bf Conceptual overview of photometric reduction pipeline.}
    Details are given in \S~\ref{sec:method}.
	\label{fig:pipeline}
	}
\end{figure}

@@

We begin with the calibrated full frame images produced by the Science
Processing Operations Center at NASA Ames (\S~\ref{subsec:observations}).
We then perform a collection of
preparatory steps, including source extraction of bright stars,
astrometric calibration, and coarse simple aperture
photometry (\S~\ref{subsec:preparation}).  Using the shape values from
the initial astrometry, we select an astrometric reference frame to
which we transform all of the calibrated images.  We construct a
photometric reference by stacking a collection of frames,
convolve the transformed frames to this photometric reference, and
subtract (\S~\ref{subsec:imagesubtraction}).  We perform aperture
photometry on the subtracted images using positions projected onto the
frame from Gaia DR2.  
We detrend the resulting light curves
(\S~\ref{subsec:lcdetrending}).  The resulting white noise and red noise properties
of the light curves, and a few
interesting cases of variability, are discussed in \S~\ref{sec:results}.

@@

@@

Having prepared the images, we perform some initial analysis steps
to produce metadata
needed during image subtraction.  To obtain an astrometric solution
(independent from the solution included with the frames), we use
\texttt{fistar} to perform source extraction on bright stars in each
image.  We pass the resulting source lists through
\texttt{astrometry.net} \citep{lang_2010}, which returns an
astrometric solution for each frame packaged in the WCS format
\citep[][Sec.~8]{pence_fits_2010}.  During the initial source
extraction, we also fit elongated gaussians to the bright stars,
yielding the shape parameters $(s,d,k)$, where the flux as a function
of position is assumed to take the form
\begin{align}
  f_{\rm elong}(\vec{x}) &= B + A \exp \{ -0.5 \times ( 
    s(\Delta x^2 + \Delta y^2) + \\
    \nonumber
    &d(\Delta x^2 - \Delta y^2) +
    k(2\Delta x \Delta y)
  )  \},
\end{align}
for $\Delta x = x-x_0$, and $\Delta y = y - y_0$.  For a nearly
circular shape profile, the sharpness $s$ is related to the FWHM as
${\rm FWHM} \approx 2.35\sqrt{s}$ \citep[{\it e.g.},][]{Pal_2009}.
These shape parameters are later used when selecting an astrometric
reference (\S~\ref{subsec:imagesubtraction}).  We note that the fast
focal ratio of the TESS cameras introduces significant comatic
aberrations to the images: stars closer to the center of the field are
more round, while stars towards the edges are more triangular.

With the resulting WCS information, we then project a source catalog
onto each frame.
Initially, we intended to photometer all Gaia-DR2 sources in each field
down to a cutoff of $G_{Rp} < 16$.
However, we found that for the galactic plane fields this produces an
excessively large number of sources.
Our source catalog for each frame is therefore a combination of
the CDIPS target stars ($G_{Rp} < 16$), and all Gaia-DR2 sources down to
$G_{Rp} < 13$.
We use the expected positions of these sources
to center the apertures in our photometry, rather than attempting to
detect the positions.  Such ``forced-aperture photometry'' is
preferable to performing source extraction simply because the TESS
pixels are large, and we are operating in crowded fields.
The Gaia-DR2 epoch is
J2015.5, so even the fastest-moving stars with proper motions of
$\sim$$1\,{\rm arcsecond}\,{\rm yr}^{-1}$ are still well within one pixel
of their predicted positions in the TESS images.  The projection from
catalog sky-coordinate positions to pixel coordinates
is performed using standard WCS transformations, 
and the source catalog look-up is
performed using
\texttt{gaia2read}\footnote{\url{github.com/samuelyeewl/gaia2read}}
\citep{kim_2018_gaia2read}.

@@

With a kernel selected, and the convolution and subtraction performed, 
we calculate the instrumental fluxes on each frame per 
Equation~\ref{eq:ism_flux_measurement}. To finally convert from a list of 
sources on each frame to a list of flux values at any given time, we use the 
\texttt{grcollect} transposition tool.

\subsection{Light curve Detrending}
\label{subsec:lcdetrending}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.6\textwidth]{f3.png}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Timeseries of ``external parameters'' for a representative
    star.} {\it Top}: Instrumental raw magnitude (with a particular
    aperture size), as a function of time.  {\it Second and third from
    top}: $x$ and $y$ centroid positions as a function of time.
    Continuing in order are the CCD temperature, the $(s,d,k)$ shape
    parameters, and the measured background value.
    Most of the apparent variability is instrumental: see
    \S~\ref{sec:flux_vs_external_parameters}.
		\label{fig:external_parameter_timeseries}
	}
\end{figure*}

\begin{figure*}[t]
    \begin{center}
		\includegraphics[height=0.9\textheight]{f4.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
      {\bf Flux as a function of ``external parameters'' for four
      representative stars.} The left two columns are stars at the
      corner of a camera's field; the right two columns are from the
      centers.
      Each row shows a different parameter along the $x$-axis, given
      in text at the bottom right of each subplot.
      ``Hooks'' are common features in flux as a function of
      temperature and centroid position.
      \S~\ref{sec:flux_vs_external_parameters} gives a verbose
      description.
     \label{fig:flux_vs_external_parameters}
    }
\end{figure*}

>>>>

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.7\textwidth]{cluster_positions.pdf}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf There are many clusters; TESS looks at them.} Placeholder skymap.
		\label{fig:clustermap}
	}
\end{figure*}

@@

\begin{figure*}[!ht]
  \gridline{\fig{astromresidual_hist.png}{0.5\textwidth}{}}
  \vspace{-0.8cm}
  \gridline{\fig{astromresidual_quiver.png}{0.5\textwidth}{}}
  \vspace{-0.8cm}
  \gridline{\fig{astromresidual_apertures.png}{0.5\textwidth}{}}
  \caption{
      {\it Top.} Histogram
      {\it Middle.} Vector
      {\it Bottom.} Apertures
  }
  \label{fig:xmatch_info}
\end{figure*}

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.3\textwidth]{pipelineoverview.pdf}
	\end{center}
	\vspace{-0.2cm}
	\caption{
    {\bf Conceptual overview of photometric reduction pipeline.}
    Details are given in \S~\ref{sec:method}.
	\label{fig:pipeline}
	}
\end{figure}

@@

We begin with the calibrated full frame images produced by the Science
Processing Operations Center at NASA Ames (\S~\ref{subsec:observations}).
We then perform a collection of
preparatory steps, including source extraction of bright stars,
astrometric verification, and coarse simple aperture
photometry (\S~\ref{subsec:preparation}).  Using the shape values from
the initial astrometry, we select an astrometric reference frame to
which we transform all of the calibrated images.  We construct a
photometric reference by stacking a collection of frames,
convolve the transformed frames to this photometric reference, and
subtract (\S~\ref{subsec:imagesubtraction}).  We perform aperture
photometry on the subtracted images using positions projected onto the
frame from Gaia DR2.  
We detrend the resulting light curves
(\S~\ref{subsec:lcdetrending}).  The resulting white noise and red noise properties
of the light curves, and a few
interesting cases of variability, are discussed in \S~\ref{sec:results}.

@@

\begin{figure*}[!ht]
  \gridline{\fig{mwscmatchstats.png}{0.95\textwidth}{}}
  \vspace{-0.8cm}
  \gridline{\fig{dias14matchstats.png}{0.95\textwidth}{}}
  \caption{
      {\it Top.} Cross-match statistics from \cite{Kharchenko_et_al_2013} {\it 
      vs.} Gaia-DR2.
      {\it Bottom.} Ditto, for \cite{dias_proper_2014} {\it vs.} Gaia-DR2.
  }
  \label{fig:xmatch_info}
\end{figure*}

@@

Having prepared the images, we perform some initial analysis steps
to produce metadata
needed during image subtraction.

First, we use \texttt{fistar} to perform source extraction on bright
stars in each image.
uring the initial source
extraction, we also fit elongated gaussians to the bright stars,
yielding the shape parameters $(s,d,k)$, where the flux as a function
of position is assumed to take the form
\begin{align}
  f_{\rm elong}(\vec{x}) &= B + A \exp \{ -0.5 \times ( 
    s(\Delta x^2 + \Delta y^2) + \\
    \nonumber
    &d(\Delta x^2 - \Delta y^2) +
    k(2\Delta x \Delta y)
  )  \},
\end{align}
for $\Delta x = x-x_0$, and $\Delta y = y - y_0$.  For a nearly
circular shape profile, the sharpness $s$ is related to the FWHM as
${\rm FWHM} \approx 2.35\sqrt{s}$ \citep[{\it e.g.},][]{Pal_2009}.
These shape parameters are later used when selecting an astrometric
reference (\S~\ref{subsec:imagesubtraction}).  We note that the fast
focal ratio of the TESS cameras introduces significant comatic
aberrations to the images: stars closer to the center of the field are
more round, while stars towards the edges are more triangular.

For the astrometric solution, we use the WCS and SIP coefficients
derived by SPOC and included in the FFI headers \citep[][Sec.~8]{pence_fits_2010}.  We
explored the possibility of using \texttt{astrometry.net}
\citep{lang_2010} to derive our own astrometric solutions for each
frame, but found that the astrometric residual (the mean separation
between projected and measured positions) was consistently a factor of
1.5-2x higher in our WCS solutions than in those given by SPOC.

With the resulting WCS information, we then project a source catalog
onto each frame.  Initially, we planned to photometer all Gaia-DR2
sources in each field down to a cutoff of $G_{Rp} < 16$.  However, we
found that for the galactic plane fields this produces an excessively
large number of sources.  Our source catalog for each frame is
therefore a combination of the CDIPS target stars ($G_{Rp} < 16$), and
all Gaia-DR2 sources down to $G_{Rp} < 13$.  We use the expected
positions of these sources to center the apertures in our photometry,
rather than attempting to detect the positions.  Such
``forced-aperture photometry'' is preferable to performing source
extraction in the crowded fields that are the focus of this work.  The
Gaia-DR2 epoch is J2015.5, so even the fastest-moving stars with
proper motions of $\sim$$1\,{\rm arcsecond}\,{\rm yr}^{-1}$ are still
well within one pixel of their predicted positions in the TESS images.
The projection from catalog sky-coordinate positions to pixel
coordinates is performed using standard WCS transformations, and the
source catalog look-up is performed using
\texttt{gaia2read}\footnote{\url{github.com/samuelyeewl/gaia2read}}
\citep{kim_2018_gaia2read}.
The astrometric residual is displayed in various forms in
Figure~\ref{fig:astrometricresidual}.

@@

With a kernel selected, and the convolution and subtraction performed, 
we calculate the instrumental fluxes on each frame per 
Equation~\ref{eq:ism_flux_measurement}. 
We do this with three different aperture sizes: for this work, circles
of radii 1 pixel, 1.5 pixels, and 2.25 pixels.
These sizes were chosen to roughly span the range of optimal aperture
sizes for stars in our sample, as calcuated in the pre-flight
\citet{Sullivan_et_al_2015} work.
To finally convert from a list of 
sources on each frame to a list of flux values at any given time, we use the 
\texttt{grcollect} transposition tool.

\subsection{Light Curve Detrending}
\label{subsec:lcdetrending}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.6\textwidth]{epdparams_vs_time.png}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Timeseries of ``external parameters'' for a representative
    star.} {\it Top}: Instrumental raw magnitude (with a particular
    aperture size), as a function of time.  {\it Second and third from
    top}: $x$ and $y$ centroid positions as a function of time.
    Continuing in order are the CCD temperature, the $(s,d,k)$ shape
    parameters, and the measured background value.
    Most of the apparent variability is instrumental: see
    \S~\ref{sec:flux_vs_external_parameters}.
		\label{fig:external_parameter_timeseries}
	}
\end{figure*}

\begin{figure*}[t]
    \begin{center}
		\includegraphics[height=0.9\textheight]{mag_vs_epdparams.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
      {\bf Flux as a function of ``external parameters'' for four
      representative stars.} The left two columns are stars at the
      corner of a camera's field; the right two columns are from the
      centers.
      Each row shows a different parameter along the $x$-axis, given
      in text at the bottom right of each subplot.
      ``Hooks'' are common features in flux as a function of
      temperature and centroid position.
      \S~\ref{sec:flux_vs_external_parameters} gives a verbose
      description.
     \label{fig:flux_vs_external_parameters}
    }
\end{figure*}
